---
title: 'Analysis for syncopation study'
author: 'nfram'
date: 'May 23, 2019'
output: html_document
---

# Preliminary work

## Clear workspace
This just clears the existing variables from the workspace.
```{r}
rm(list=ls())
```

## Read in necessary libraries

```{r}
library(lme4)
library(lmerTest)
library(languageR)
library(tidyverse)
```

## Set proper directory
Note: make sure that this actually points to the right file path! It's a bit fragile and hacky...
```{r}
setwd('~/Documents/Research/Linguist_245b/my_project/analysis/01_experiment/rscripts/')
```

# Data import and rearrangement

## Read in datasets
Just import all the necessary data.

```{r}
sync_data = read.csv('../data/syncopation-trials.csv')
background_data = read.csv('../data/syncopation-background_information.csv')
worker_data = read.csv('../data/syncopation-subject_information.csv')
head(sync_data)
head(background_data)
```

And parse for experimental trials, not attention checks.
```{r}
sync_meas = inner_join(sync_data, background_data, by=c('workerid')) # Add background data by subject
head(sync_meas)
```

## Parse to only real trials, not attention check
```{r}
test_sync_d = sync_meas[sync_meas$fac1 != -99,]
catch_sync_data = sync_meas[sync_meas$fac1 == -99,]
head(test_sync_d)
head(catch_sync_data)
ggplot(data = test_sync_d, aes(x = workerid, y = measured_sync)) +
  geom_point(color = 'blue') +
  geom_point(data = catch_sync_data, color = 'red')
```

## Exclude with weird attention check rhythms
```{r}
tsd2 = test_sync_d %>%
  group_by(workerid)
csd2 = catch_sync_data %>%
  group_by(workerid)
csd2 %>%
  summarize(max_catch = max(measured_sync))
test_mins = tsd2 %>%
  summarize(min_test = min(measured_sync))
head(test_mins)
catch_update = left_join(csd2, test_mins)
head(catch_update)
c_u <- catch_update %>%
  ungroup()
c_u$measured_sync
c_u$min_test
c_u_v2 <- c_u %>%
  mutate(Is_Valid = if_else(measured_sync <= min_test, 0, 1))
head(c_u_v2)
c_u_group <- c_u_v2 %>%
  group_by(workerid)
Num_Invalid <- c_u_group %>%
  summarize(num_invalid = sum(Is_Valid))
head(Num_Invalid)
catch_groups <- left_join(tsd2, Num_Invalid)
head(catch_groups)
test_sync_data <- catch_groups[catch_groups$num_invalid < 2,]

```
## Convert doesplay, perftime, playtime, and listenfreq to categorical variables
```{r}
test_sync_data <- test_sync_data %>%
  mutate(doesplay = recode(doesplay, '0'='No', '1'='Yes')) %>%
  mutate(listenfreq = recode(listenfreq, '0' = 'Never', '1' = 'Rarely', '2'='Sometimes', '3'='Often', '4'='Daily')) %>%
  mutate(perftime = recode(perftime, '0' = 'Never', '1' = 'Rarely', '2'='Sometimes', '3'='Often', '4'='Daily')) %>%
  mutate(playtime = recode(playtime, '0' = 'Less than 6 months', '1' = '6-12 months', '2' = '1-2 years', '3' = '2-5 years', '4' = 'More than 5 years')) %>%
  mutate(practime = recode(practime, '0' = 'Never', '1' = 'Rarely', '2'='Sometimes', '3'='Often', '4'='Daily'))
```

# Statistical analysis

## Demographics
Compute basic demographic statistics for:
age | gender | doesPlay | playTime | listenType
```{r}
age_mean <- mean(worker_data$age)
age_sd <- sd(worker_data$age)

summary(worker_data$gender)

summary(worker_data$fairprice)

summary(background_data$instname)

summary(worker_data)
summary(background_data$doesplay)
background_data %>%
  filter(doesplay == 0) %>%
  count()

summary(background_data$listenfreq)
freqtable_listen <- table(background_data$listenfreq) %>%
  as.data.frame.table()
head(freqtable_listen)

freqtable <- table(background_data$listentype) %>%
  as.data.frame.table()
head(freqtable)

ggplot(data = freqtable, aes(Var1, Freq)) +
  geom_bar(stat='identity')

ggplot(data = background_data, aes(fill=listentype)) +
  geom_histogram()

```

## Univariate correlations
These mimic the analysis performed by Smith and Honing (2006)
```{r}
m_lhlfitch = lm(measured_sync ~ lhl.fitch, data = test_sync_data)
summary(m_lhlfitch)
m_metcomp = lm(measured_sync ~ metComp, data = test_sync_data)
summary(m_metcomp)
m_wnbd1 = lm(measured_sync ~ wnbd1, data = test_sync_data)
summary(m_wnbd1)

ggplot(test_sync_data, aes(x=lhl.fitch, y=measured_sync)) +
  geom_point() +
  geom_smooth(method = lm) +
  ylab('Measured syncopation rating') +
  xlab('Longuet-Higgins and Lee syncopation')
ggsave('lhl_sync_corr.png')

ggplot(test_sync_data, aes(x=metComp, y=measured_sync)) +
  geom_point() +
  geom_smooth(method = lm) +
  ylab('Measured syncopation rating') +
  xlab('Metrical complexity')
ggsave('metcomp_sync_corr.png')

ggplot(test_sync_data, aes(x=wnbd1, y=measured_sync)) +
  geom_point() +
  geom_smooth(method = lm) +
  ylab('Measured syncopation rating') +
  xlab('Weighted note-to-beat distance (whole note)')
ggsave('wnbd1_sync_corr.png')
  
```

## Mixed models
Mixed model using factor 1, factor 2, and conditional intercepts for play time, listening frequency, and performance time (as indices of musical engagement).
```{r}
m = lmer(measured_sync ~ (1|workerid) + (fac1 + fac2)*(doesplay + playtime + listenfreq + practime), data = test_sync_data, REML = F, control=lmerControl('Nelder_Mead'))
summary(m)
```

And single-variable versions
```{r}
m1 = lmer(measured_sync ~ (1|workerid) + (fac1)*(doesplay + playtime + listenfreq), data = test_sync_data, REML = F, control=lmerControl('Nelder_Mead'))
summary(m1)

m2 = lmer(measured_sync ~ (1|workerid) + (fac2)*(doesplay + playtime + listenfreq), data = test_sync_data, REML = F, control=lmerControl('Nelder_Mead'))
summary(m2)
```

## Structural equation model

# Visualizations
```{r}
new_sync_data = test_sync_data %>%
  mutate(dp_cat = recode(doesplay, '0'='No', '1'='Yes'))

listen_sync_data = new_sync_data %>%
  mutate(lf_cat = recode(listenfreq, '0' = 'Never', '1' = 'Rarely', '2'='Sometimes', '3'='Often', '4'='Daily'))

summary(listen_sync_data$lf_cat)

new_fac_data = new_sync_data %>%
  gather(WhichFactor, FactorValue, fac1, fac2)
head(new_fac_data)

listen_fac_data = listen_sync_data %>%
  gather(WhichFactor, FactorValue, fac1, fac2)
head(listen_fac_data)

vars <- c('Factor 1'='red', 'Factor2'='blue')

p1 <- ggplot(subset(listen_fac_data, !is.na(lf_cat)), aes(x=FactorValue, y=measured_sync, color=WhichFactor)) +
  geom_smooth(method = 'lm') +
  facet_wrap(~lf_cat, labeller = label_bquote('Listens to music' : .(lf_cat))) +
  xlab('Average z-score for each factor') +
  ylab('Measured syncopation rating') +
  labs(color = 'Component') +
  scale_color_discrete(breaks = c('fac1','fac2'), labels = c('Density','Hierarchy'))

p1

ggsave('corr_by_listenfreq.png', width=8.1, height=5, units='in')

ggsave('corr_by_doesplay.png')

ggplot(subset(new_sync_data, !is.na(dp_cat)), aes(x=fac2, y=measured_sync, color=dp_cat)) +
  geom_smooth(method='lm')

ggplot(subset(test_sync_data, !is.na(doesplay)), aes(x=fac2, y=measured_sync)) +
  geom_smooth(method='lm') +
  facet_wrap(~doesplay)
```

```{r}
ggplot(test_sync_data)
```